Wildcard: Spreadsheet-Driven Customization of Web Applications

### 2020-01-14 paper.tex
submission ready

The first party developers of the original software can also play a role
in enabling customization. Although spreadsheet-driven customization
does not depend on cooperation from first-party website developers,
their participation in exposing structured data would eliminate the need
for third party site adapters. The first parties could themselves
benefit from such a system, since providing Wildcard support---a fairly
low-effort undertaking for a developer with direct access to the
structured data---would allow users to build extensions to fulfill some
of their own feature requests. There is also precedent for first parties
implementing an official client extension API: for several years, Google
maintained an official extension API in Gmail for Greasemonkey scripts
to use.

@@

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. In addition to changes to the website,
  changes to the system's heuristics can also cause unexpected changes
  in behavior.
\item
  \emph{Discoverability}: it can be hard for users to find discover
  possible commands.
\end{itemize}

@@

Another difference between Wildcard and these tools is that there is no
way for users to express imperative workflows with sequences of actions
in Wildcard. Chickenfoot and Coscripter have shown these types of
workflows to be useful in practice so it might be worth considering how
to support them, but it's not clear how such workflows could fit into
Wildcard. Spreadsheets have a fundamentally different computation model,
and site adapters created by programmers cannot easily account for every
possible action a user might want to perform in a page.

@@

Many researchers and companies have explored the idea of
\emph{spreadsheet-driven applications}: using spreadsheets as a backing
data store and computation layer for interactive web applications,
enabling end users to create such applications more easily. Research
projects like Object Spreadsheets \autocite{mccutchen2016}, Quilt
\autocite{benson2014}, Gneiss \autocite{chang2014}, and Marmite
\autocite{wong2007}, as well as commercial tools like Airtable
\autocite{zotero-79} and Glide \autocite{zotero-81} allow users to view
data in a spreadsheet table, compute over the data using formulas, and
then connect the table to a GUI. Because many users are already familiar
with using spreadsheets, this way of creating applications tends to be
much easier than traditional software methods; for example, in a user
study of Quilt, many users were able to create applications in under 10
minutes, even if they expected it would take them many hours.

@@

We hope that spreadsheet-driven customization can help the Web reach its
full potential as a dynamic medium that users can mold to their own
needs.

\acks
Thanks to Glen Chiacchieri, David Karger, Steve Krouse, Rob Miller, Arvind Satyanarayan, Daniel Windham, and Maggie Yellen for providing valuable feedback on this work.

>>>>

The first party developers of the original software can also play a role
in enabling customization. Although spreadsheet-driven customization
does not depend on cooperation from first party website developers,
their participation in exposing structured data would eliminate the need
for third party site adapters. The first parties could themselves
benefit from such a system; providing Wildcard support---a fairly
low-effort undertaking for a developer with direct access to the
structured data---would allow users to build extensions to fulfill some
of their own feature requests. There is also precedent for first parties
implementing an official client extension API: for several years, Google
maintained an official extension API in Gmail for Greasemonkey scripts
to use.

@@

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. Changes to the website or to the system's
  heuristics can cause unexpected changes in behavior.
\item
  \emph{Discoverability}: it can be difficult for users to find explore
  the space of possible commands.
\end{itemize}

@@

Another difference between Wildcard and these tools is that there is no
way for users to express imperative workflows with sequences of actions
in Wildcard. Chickenfoot and Coscripter have shown these types of
workflows to be useful in practice so it might be worth considering how
to incorporate them, but it's not obvious how such workflows could fit
into Wildcard. Spreadsheets have a fundamentally different computation
model, and site adapters created by programmers cannot easily account
for every possible action a user might want to perform in a page.

@@

Many researchers and companies have explored the idea of
\emph{spreadsheet-driven applications}: using spreadsheets as a backing
data store and computation layer for interactive web applications,
enabling end users to create such applications more easily. Research
projects like Object Spreadsheets \autocite{mccutchen2016}, Quilt
\autocite{benson2014}, Gneiss \autocite{chang2014}, and Marmite
\autocite{wong2007}, as well as commercial tools like Airtable
\autocite{zotero-79} and Glide \autocite{zotero-81} allow users to view
data in a spreadsheet table, compute over the data using formulas, and
then connect the table to a GUI. Because many users are already familiar
with spreadsheets, this way of creating applications tends to be much
easier than traditional software methods; for example, in a user study
of Quilt, many users were able to create applications in under 10
minutes, even if they expected it would take them many hours.

@@

Ultimately, we hope that spreadsheet-driven customization helps the Web
reach its full potential as a dynamic medium that users can mold to
their own needs.

\acks
Thanks to Tarfah Alrashed, Glen Chiacchieri, David Karger, Steve Krouse, Rob Miller, Santiago Perez De Rosso, Arvind Satyanarayan, Daniel Windham, and Maggie Yellen for providing valuable feedback on this work.
